---
title: "Naomi Struble, nas4252"
output: html_document
date: "2024-02-15"
---

Github link:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
ebay <- read.csv('ebay.csv')
gasprices <- read.csv('gasprices.csv')
nbc <- read.csv('nbc_pilotsurvey.csv')
cars <- read.csv('sclass.csv')
library(tidyverse)
library(mosaic)
```

------------------------------------------------------------------------

## **Problem 1**
#### Theory A: 
```{r, echo=FALSE, results='hide'}
mean(Price ~ Competitors, data=gasprices)
boot_competitors = do(10000)*diffmean(Price ~ Competitors, data=mosaic::resample(gasprices))
confint(boot_competitors, level = 0.95)
```

**Claim:** Theory A claims that gas stations charge more if they lack direct competition in sight.

**Evidence:** The differences in means between gas stations that do have direct competition in sight and those that don't, show that those that don't have competition in sight charge more on average. Furthermore, the difference in price between gas stations on and off the highway is somewhere between -.05 and .01, with 95% confidence.

**Conclusion:** Based on the differences in means it could be said that the theory is correct; however, because the results of the confidence interval are statistically insignificant, there is no statistical evidence that the theory is true.

#### Theory B:
```{r, echo=FALSE, results='hide'}
options(scipen = 999)
mean(Price ~ Income, data=gasprices)
boot_income = do(10000)*lm(Price ~ Income, data=mosaic::resample(gasprices))
confint(boot_income, level = 0.95)
```
```{r, echo=FALSE, message=FALSE}
ggplot(gasprices) + 
  geom_jitter(aes(x=Price, y=Income), alpha=0.1) + 
  geom_smooth(aes(x=Price, y=Income), method='lm')
```

**Claim:** Theory B claims that the richer the area, the higher the gas prices.

**Evidence:** The difference in price between gas stations in rich areas and those outside rich areas is somewhere between .0000007 and .000002, with 95% confidence. Additionally, using the scatter plot graph above, there is further evidence that the price of gas increases as income increases.

**Conclusion:** Based on the evidence, it can be concluded that the theory is true. In fact, the results are statistically significant based on the confidence interval.

#### Theory C:
```{r, echo=FALSE, results='hide'}
mean(Price ~ Stoplight, data=gasprices)
boot_stoplight = do(10000)*diffmean(Price ~ Stoplight, data=mosaic::resample(gasprices))
confint(boot_stoplight, level = 0.95)
```

**Claim:** Theory C states that gas stations at stoplights charge more.

**Evidence:** When analyzing this statement, it is important to first analyze the mean prices for gas stations at stoplights and those that are not. In doing so, the average gas price for gas stations at stoplights is lower than those not at stoplights. Additionally, the difference in price between gas stations at stoplights and those that aren't is somewhere between -.04 and .03, with 95% confidence.

**Conclusion:** Based on this evidence, it can be concluded that the theory is false, for prices tend to be lower at gas stations at stoplights. Furthermore, the confidence interval reveals that the findings are statistically insignificant because zero is included in the interval.

#### Thoery D:
```{r, echo=FALSE, results='hide'}
mean(Price ~ Highway, data=gasprices)
boot_highway = do(10000)*diffmean(Price ~ Highway, data=mosaic::resample(gasprices))
confint(boot_highway, level = 0.95)
```

**Claim:** Theory D claims that gas stations with direct highway access charge more.

**Evidence:** To further analyze the claim, it is important to calculate the average gas prices for gas stations that have direct access to the highway and those that don't. In doing so, it is shown that the gas stations with direct highway access do, in fact, charge more. Additionally, the difference in price between gas stations with direct highway access and those without is somewhere between .009 and .081, with 95% confidence.

**Conclusion:** As a result of the evidence, it can be concluded that the theory is true. In fact, the confidence interval is proven to be statistically significant, further amplifying the theory.

#### Theory E:
```{r, echo=FALSE, results='hide'}
gas_shell = gasprices %>%
  mutate(shell_or_not = ifelse(Brand == 'Shell', yes = 'Y', no = 'N'))
mean(Price ~ shell_or_not, data=gas_shell)
boot_brand = do(10000)*diffmean(Price ~ shell_or_not, data=mosaic::resample(gas_shell))
confint(boot_brand, level = 0.95)
```

**Claim:** Theory E states that Shell gas stations charge more than all other non-Shell brands.

**Evidence:** In order to analyze the claim, it is important to first calculate the average prices of gas stations that are Shell and those that are other brands. After making these calculations, it is shown that Shell brands do have higher gas prices on average. Additionally, the difference in price between gas stations with direct highway access and those without is somewhere between -.001 and .064, with 95% confidence.

**Conclusion:** Although the calculated means from the sample depict the idea that Shell gas stations charge more than the other non-Shell brands, the statistically insignificant results from the confidence interval negate these findings. As a result it can be concluded that the theory is false.

------------------------------------------------------------------------

## **Problem 2**
#### Part A:
```{r, echo=FALSE}
class <- cars %>%
  filter(year == 2011, trim == "63 AMG")

boot_class = do(10000)*mean(~ mileage, data=mosaic::resample(class))
confint(boot_class, level = 0.95)
```

#### Part B:
```{r, echo=FALSE}
blackclass <- cars %>%
  filter(year == 2014, trim == "550") %>%
  mutate(isBlack = ifelse(color == "Black", yes=TRUE, no=FALSE))

boot_blackclass = do(10000)*prop(~ mileage, data=mosaic::resample(blackclass))
confint(boot_blackclass, level = 0.95)
```

------------------------------------------------------------------------

## **Problem 3**
#### Part A:

**Question:** Is there evidence that one show consistently produces a higher mean Q1_Happy response among viewers?

**Approach:** The first step I took was filtering the original data set to only list the shows " Living with Ed" and "My Name is Earl". I then calculated the difference in means between the happiness ratings for the shows "Living with Ed" and "My Name is Earl". And then I took a bootstrap of the difference in mean among 10,000 samples.
```{r, echo=FALSE}
nbc_happy <- nbc %>%
  filter(Show == "Living with Ed" | Show == "My Name is Earl")

boot_nbc_happpy = do(10000)*diffmean(Q1_Happy ~ Show, data=mosaic::resample(nbc_happy))
```

**Results:** The results from the difference in mean from the original sample is as follows:

```{r, echo=FALSE}
mean(Q1_Happy ~ Show, data=nbc_happy)
```

This demonstrates that the average rating for "Living with Ed" is higher than "My Name is Earl" in the original sample. Furthermore, by taking the confidence interval from the bootstrap, more information can be gathered. The difference in happiness response among viewers between "Living with Ed" and "My Name is Earl" is somewhere between -.399 and .096, with 95% confidence.

```{r, echo=FALSE, results='hide'}
confint(boot_nbc_happpy, level = 0.95)
```

**Conclusion:** Based on the results of the means calculated, it does appear that "Living with Ed" produces a happier response rating among its viewers. However, the confidence interval demonstrates that by taking the difference in means among thousands of samples, the results become statistically insignificant. Therefore, it can be concluded that neither show consistently produces a higher mean happiness response among viewers.

#### Part B:

**Question:** Is there evidence that one show consistently produces a higher mean Q1_Annoyed response among viewers?

**Approach:** The First step I took to analyze the data set was filtering the sample provided so that only the shows "The Apprentice: Los Angeles" and "The Biggest Loser" were included. Next, I calculated the difference in means between the annoyed responses for each of the shows. Lastly, I took a bootstrap of the difference in means among 10,000 samples of the shows.
```{r, echo=FALSE}
nbc_annoyed <- nbc %>%
  filter(Show == "The Biggest Loser" | Show == "The Apprentice: Los Angeles")

boot_annoyed = do(10000)*diffmean(Q1_Annoyed ~ Show, data=mosaic::resample(nbc_annoyed))
```

**Results:** When calculating the difference in means of the annoyed response for the two shows, the results are as follows:

```{r, echo=FALSE}
mean(Q1_Annoyed ~ Show, data=nbc_annoyed)
```

The results above demonstrated how the show "The Apprentice: Los Angeles" receives a higher average annoyed response. Additionally, I calculated the confidence interval from the bootstrap I conducted earlier.

```{r, echo=FALSE, results='hide'}
confint(boot_annoyed, level = .95)
```

The difference in happiness response among viewers between "The Apprentice: Los Angeles" and "The Biggest Loser" is somewhere between -.519 and -.020, with 95% confidence.

**Conclusion:** Based on the results of the means calculated, it does appear that "The Apprentice: Los Angeles" produces a more annoyed response rating among its viewers. Additionally, the confidence interval demonstrates that by taking the difference in means among thousands of samples, the results become statistically significant. Therefore, it can be concluded that "The Apprentice: Los Angeles" does in fact consistently produces a higher mean annoyed response among viewers.

#### Part C:

**Question:** What proportion of American TV watchers would we expect to give a response of 4 or greater to the "Q2_Confusing" question?

**Approach:** The Approach I took when attempting to answer the question was first filtering the data set to only include the show "Dancing with the Stars". I then mutated the new data set so that a new column was created that would tell me if there was a rating greater than or equal to four. Lastly, I bootstrapped the new data set by the proportion of viewers that rated the show a four or five in being confusing.

```{r, echo=FALSE}
nbc_dancing <- filter(nbc, Show == "Dancing with the Stars")
nbc_dancing <- mutate(nbc_dancing, nbc_confusing = ifelse(Q2_Confusing >= 4, "Yes", "No"))

boot_confusing <- do(10000) * prop(~nbc_confusing, data = mosaic::resample(nbc_dancing))
```

**Results:** After calculating the bootstrap, I then created a confidence interval that depicts the proportion of responses equal to or greater than four is between .044 and .116, with 95% confidence.
```{r, echo=FALSE, results='hide'}
confint(boot_confusing, level = .95)
```

```{r, echo=FALSE, results='hide'}
mean(nbc_dancing$Q2_Confusing >= 4)
```
**Conclusion:** In order to approximate the proportion of responses that rated "Dancing with the Stars" a 4/5 or greater based on how confusing it is, it is important to analyze both the original data set and its bootsrap samples. According to the data sample, the proportion of responses that rated the show a 4 or above is `r round(mean(nbc_dancing$Q2_Confusing >= 4), digits=3)`. While this number does accurately represent the original sample, the confidence interval demonstrates that there is a variance in that number, depending on the sample. 

------------------------------------------------------------------------

## **Problem 4**

**Question:** Is the revenue ration the same in the treatment and control groups or does paid search advertising on Google create extra revenue for EBay?

**Approach:** The approach I took first was filtering the original data set to separate the treatment and control groups. This way I was able to calculate the difference in proportion of the revenue ratio for the two groups. Afterwards, I conducted a bootstrap of 10,000 samples.

```{r, echo=FALSE, results='hide'}
#Filter control and treatment
ebay_0 <- ebay %>%
  filter(adwords_pause == "0")
ebay_1 <- ebay %>%
  filter(adwords_pause == "1")

#Determine prop of control and treatment
prop(ebay_0$rev_ratio)
prop(ebay_1$rev_ratio)

#Bootstrap
boot_ebay = do(10000)*diffprop(rev_ratio ~ adwords_pause, data=mosaic::resample(ebay))
```

**Results:** After calculating the proportion of revenue that was contributed by the control(.006) and treatment groups(.018), there is a significantly higher proportion of revenue contributed by the treatment group. Additionally, the results from the confidence interval show that there is a difference in the proportion of revenue gained by the control group and treatment group between -.013 and .064, with 95% confidence.
```{r, echo=FALSE, results='hide'}
confint(boot_ebay, level = 0.95)
```


**Conclusion:** Although the difference in proportion may signal that paid search advertising on Google does generate extra revenue for EBay, the statistically insignificant result from the confidence interval prove the theory to be incorrect. 